DMRGenie: Fully graphical user interface for entanglement renormalization

### 2024-12-28 paper/paper.tex
Add files via upload

\usepackage{amsmath}

@@

We discuss implementation of a graphical user interface for tensor operations in scientific simulations related to entanglement renormalization. This software, which we call DMRGenie, based on its heavy reliance on Genie which is a well-known software package for web applications. We also discuss the deployment of the software. Scientific simulations with tensor network-based framework of quantum algorithm methods is now possible through this tool without length start-up or learning times.

@@

We will first discuss the background software (DMRjulia and TENPACK) originally developed in 2016 and first introduced to Github in 2020 and full debut on the Julia package environment in 2021. This means that the library has spanned many version of the Julia version starting originally in v0.5 and now continuing to v1.11 and beyond. We believe that some choices from our implementation may be interesting to the broader community, although many of the improvements that were introduced here later appeared in this development independently appeared in Julia from other groups, showing that many issues are being encountered in common and handled in similar ways.

@@

The issue of type stability has been of paramount concern to ensure that any code developed in the language avoids generating too much overhead because of Julia's multiple dispatch capabilities. Because the compiler in Julia must assume that any possible input can be entered into a function, the assumptions of how types are converted and exactly what type is created can generate excess allocations which ultimately slow a program. Eliminating these excess allocations is necessary to create efficient code that is competitive with lower level languages.

@@

The original hope for Julia was to solve the two-language problem, and that is still largely solved in DMRjulia and TENPACK. We have merely made some compromise choices to ensure that there is efficient implementation because our use case is narrower than the most broad implementation that Julia strives to implement.

@@

\section{Graphical interface with DMRGenie}
The DMRGenie graphical user interface (GUI) allows for tensor network algorithms to be run with minimal to no coding required, and using diagrams to denote the algorithms that have been selected. DMRGenie consists of the Algorithm Runner and the Tensor Network Builder. The Algorithm Runner runs tensor network algorithms such as DMRG using known models or custom Hamiltonians. The Tensor Network Builder allows the user to build their own tensor network and create their own tensor network algorithm using a drag-and-drop system. The frontend was built using HTML, JavaScript, React, and CSS for styling. The backend uses a Model View Controller (MVC) infrastructure. The packages used run the tensor network functions are DMRjulia and TENPACK. So, while both DMRjulia and TENPACK are 100\% from-scratch Julia implementations, DMRGenie is not, but this makes the code easier to use and work with and adapt to changing software standards in the future.

\subsection{DMRGenie Backend}\label{backend}
The DMRGenie website was built using a Model View Controller (MVC) infrastructure.  The benefit of a MVC framework is that the backend and frontend can be worked on simultaneously which allows for smoother development. The MVC was created using the Genie.jl Julia package. Genie was chosen because it allowed for full-stack development with Julia as the backend. With Julia as the backend, it is simple to run tensor network functions with DMRjulia and TENPACK. The model and controls are written in Julia while the view is written in Julia and HTML using a ".jl.html" file extension.

The model is like an object that is passed between the frontend and the backend of the application. It is used to pass data between the two ends. When initializing the model, any variables to pass data are initialized to default values. The data returned by the model is mostly raw data.

The view file contains the HTML for the application. This file is capable of containing Julia code due to its ".jl.html" file extension. This also allows the controller to pass Julia variables to be displayed or output on the frontend. Despite the allowance of Julia code being executed in the view component, Julia was only used to present the data rather than run computations.

The controller is responsible for starting up the HTML page with the relevant data when website is initialized. When a tensor network is created or a tensor network function is run, the controller is responsible to create the tensors and run the algorithm using DMRjulia and TENPACK. The results of any operation are stored in variables and returned to the view component.

@@

It is also possible to compute the correlations and impose symmetries on the system. The correlation matrix operators and correlation function operators must be specified to compute correlations. The matrix operators and function operators will then be evaluated at all position on the lattice.

@@

\subsection{Tensor Network Builder}
The Tensor Network Builder is a drag and drop interface to build tensor networks. The rank of any tensor is specified before they are dropped onto the canvas. The number of connections points on a tensor is equal to the rank of the tensor. Indices can be created by dragging from a connection point on a tensor. It is possible to specify the name and dimension of all indices, and it is possible to rename an index if needed. Possible operations on tensors include contracting, singular value decompositions, eigenvalue decomposition, QR decompositions, and LQ decompositions. When the user is done the algorithm then the "submit" button can be clicked and the algorithm will be run on the backend. For more information about the backend, see Sec.~\ref{backend}. The output of the algorithm will be the tensors that resulted from the last operation. The output is given as a downloadable file.\\

ReactJS is used for the tensor network builder because of component reusability and the fast rendering. Many components can be reused because all the tensors and indices objects are reused. The ReactFlow package was designed to build node-based editors and interactive diagrams; however, graphs in computer science have similar structures to tensor networks. The nodes and edges in ReactFlow were generalized to tensors and indices.

>>>>

%\usepackage{amsmath}

\usepackage{amssymb,amsmath}
%\usepackage[pdftex]{graphicx}
%\usepackage[usenames,dvipsnames]{color}
%\usepackage{epstopdf}
%\usepackage[hidelinks]{hyperref}
%\usepackage{courier}
%\usepackage{subcaption}

\usepackage{subfigure}

@@

We discuss implementation of a graphical user interface for tensor operations in scientific simulations related to entanglement renormalization. We call this software DMRGenie based on its heavy reliance on Genie which is a well-known software package for web applications. We also discuss the deployment of the software. Scientific simulations with tensor network-based framework of quantum algorithm methods are now possible through this tool without lengthy start-up or learning times.

@@

We will first discuss the background software (DMRjulia and TENPACK; available in Julia's package system as {\tt DMRJtensor} and {\tt TensorPACK}) originally developed in 2016 and first introduced to Github in 2020 and full debut on the Julia package environment in 2021. This means that the library has spanned many version of the Julia version starting originally in v0.5 and now continuing to v1.11 and beyond. We believe that some choices from our implementation may be interesting to the broader community, although many of the improvements that were introduced here later appeared in this development independently appeared in Julia from other groups, showing that common issues are being encountered and handled in similar ways.

@@

The issue of type stability has been of paramount concern to ensure that any code developed in the language avoids generating too much overhead because of Julia's multiple dispatch capabilities. Since the compiler in Julia must assume that any possible input can be entered into a function, the assumptions of how types are converted and exactly what type is created can generate excess allocations which ultimately slow a program. Eliminating these excess allocations is necessary to create efficient code that is competitive with lower level languages.

@@

The original hope for Julia was to solve the two-language problem, and that is still largely solved in DMRjulia and TENPACK. We have merely made some compromises to ensure that there is efficient implementation because our use case is narrower than the most broad implementation that Julia strives to implement.

@@

\section{Graphical interface with DMRGenie}
The DMRGenie graphical user interface (GUI) allows for tensor network algorithms to be run with minimal to no coding required, and using diagrams which are easier to read. DMRGenie consists of the Algorithm Runner and the Tensor Network Builder. The Algorithm Runner executes tensor network algorithms such as DMRG using known models or custom Hamiltonians. The Tensor Network Builder allows the user to build their own tensor network and create their own tensor network algorithm using a drag-and-drop system. The frontend was built using HTML, JavaScript, React, and CSS for styling. The backend uses a Model View Controller (MVC) infrastructure. The packages used to run the tensor network algorithms are DMRjulia and TENPACK. So, while both DMRjulia and TENPACK are 100\% from-scratch Julia implementations, DMRGenie is not, but this makes the code easier to use and work with and adapt to changing software standards in the future.

\subsection{DMRGenie Backend}\label{backend}
The DMRGenie website was built using a Model View Controller (MVC) infrastructure, which was initialized using the Genie.jl Julia package. Genie was chosen because it allowed for full-stack development with Julia as the backend. This improved the development cycle as it made it simple to run tensor network functions with DMRjulia and TENPACK. The models and controls are written in Julia, while the views are written in Julia and HTML using a \texttt{.jl.html} file extension with supporting JavaScript files for functionality.

The model is a struct which is used to pass data between the frontend and backend of the application. 
When a page is initially loaded, a model is initialized with a default set of data. Once the user inputs their options and submits, that raw data is sent to the backend where tensor network computations can then be performed accordingly.

The view file contains the HTML for the application. This file is capable of containing Julia code due to its ".jl.html" file extension, allowing the controller to pass Julia variables to be output on the frontend.

The controller is responsible for starting up the HTML page with the relevant data when the website is initialized. When a tensor network is created or a tensor network function is run, the controller is responsible for creating the tensors and running the algorithm using DMRjulia and TENPACK. The results of any operation are stored in variables and returned to the view component.

\begin{figure}[b]
    \includegraphics[width=\columnwidth]{contraction.png}
\caption{Contraction two tensors with the Tensor Network Builder interface. a) The initial tensors. b) The input given to the contract menu operation. c) The output tensor. } \label{GUIcontraction}
\end{figure}

@@

It is also possible to compute the correlations and impose symmetries on the system. The correlation matrix operators and correlation function operators must be specified to compute correlations. The matrix operators and function operators will then be evaluated at all positions on the lattice.

@@

\subsection{Tensor Network Builder}
The Tensor Network Builder is a drag and drop interface to build tensor networks. The rank of any tensor is specified before they are dropped onto the canvas. The number of connections points on a tensor is equal to the rank of the tensor. Indices can be created by dragging from a connection point on a tensor. It is possible to specify the name and dimension of all indices, and it is possible to rename an index if needed. Possible operations on tensors include contractions (see example of interface in Fig.~\ref{GUIcontraction}), singular value decompositions, eigenvalue decompositions, QR decompositions, and LQ decompositions. When the user is done, the "submit" button can be clicked and the algorithm will be run on the backend. The output of the algorithm will be the tensors that resulted from the last operation. The output is given as a downloadable file.\\

ReactJS is used for the tensor network builder because of component reusability and the fast rendering. The ReactFlow package was designed to build node-based editors and interactive diagrams; however, graphs in computer science have similar structures to tensor networks. The nodes and edges in ReactFlow were generalized to tensors and indices.

For example, if the user wanted to input a Heisenberg Hamiltonian of 
\begin{equation}\label{heisMPO}
\hat H = -J\sum_{i} \left( \sigma^x_i \sigma^x_{i+1} + \sigma^y_i \sigma^y_{i+1} + \sigma^z_i \sigma^z_{i+1}\right)
\end{equation}
where $J$ is the Hamiltonian constant, then it would be possible to do so term by term. An example of the input is given in Fig.~\ref{heismodel}.

\begin{figure}
\begin{center}
    \includegraphics[scale=0.5]{HeisenbergModel.png}
\end{center}
    \caption{The Heisenberg Model input for the Algorithm Runner written term-by-term as in Eq.~\eqref{heisMPO}.
    \label{heismodel}
    }
\end{figure}
